---
title: "Lab 1: Intro to R"
author: "Leticia Salazar"
date: "September 4, 2021"
output:
  html_document: default
  pdf_document: default
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1

Counts of girls baptized:

```{r view-girls-counts}
arbuthnot$girls
```


### Exercise 2

There is a peak around 1660 with the number of girls being baptized increases with a slight decrease after the 1700s. Overall, it's an upward trend in girls being baptized.

```{r trend-girls}
ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_line()

ggplot(data = arbuthnot, aes(x = year, y = girls)) +
  geom_point()
```

### Exercise 3

The boys born over time are slightly above 50% up until early 1700s where there's a sudden drop. Overall, the birth of boys has significantly increased.

```{r plot-prop-boys-arbuthnot}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)

ggplot(data = arbuthnot, aes(x = year, y = boys/total)) + 
  geom_line()
```


### Exercise 4

The years included in this data set are from 1940 to 2002. The dimensions of the data set are 63, 3 and the column names are "year", "boys", "girls".

```{r dim-present}
data(present)

dim(present)

range(present$year)

names(present)
```


### Exercise 5

Comparing the arbuthnot's and present data set they are completely different. The median for arbuthnot's data set is 11813 where as the one for present is 3756547.

```{r count-compare}
summary(arbuthnot)

summary(present)
```


### Exercise 6

Looking at the graph, there is a decrease in boys being born in the US compared to Arbuthnot's observation. You can see that between 1940 and 1945 there's still a peak but after 1950's there's a decline. From 1965 to 1980 there's a slight increase with a huge drop afterwards.

```{r plot-prop-boys-present}
present <- present %>%
  mutate(total = boys + girls)

ggplot(data = present, aes(x = year, y = boys/total)) + 
  geom_line()
```


### Exercise 7

Based on this data, the most number of births in the US was in 1961.

```{r find-max-total}
present %>%
  mutate(total = boys + girls) %>%
  arrange(desc(total))
```

