---
title: DATA 606 Data Project Proposal
author: "Leticia Salazar"
date: "`r Sys.Date()`"
output: 
 html_document:
    theme: cosmo
---

### Data Preparation

```{r}
# load packages / libraries

#These packages are necessary for maps
#install.packages("cowplot")
#install.packages("googleway")
#install.packages("ggplot2")
#install.packages("ggrepel")
#install.packages("ggspatial")
#install.packages("libwgeom")
#install.packages("sf")
#install.packages("rnautralearth")
#install.packages("rnautralearthdata")

library("dplyr")
library("tidyverse")

#These libraries are for maps
library("ggplot2")
theme_set(theme_bw())
library("sf")

```


```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data
drinks <- read.csv("https://raw.githubusercontent.com/letisalba/Data-606/main/Project/drinks.csv", header = TRUE, na.strings = c(""," ", "NA"));
drinks
```

```{r}
glimpse(drinks)
```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

_What country has the highest alcohol consumption?_

### Cases 

**What are the cases, and how many are there?**

_Each case represents a country around the world along with their beer, spirits and/or wine number of servings, as well as the total liters of pure alcohol. There are 193 total observations in this data set._

### Data collection 

**Describe the method of data collection.**

_The data was collected from [FiveThirtyEight's](https://fivethirtyeight.com/features/dear-mona-followup-where-do-people-drink-the-most-beer-wine-and-spirits/) article called "Dear Mona Followup: Where Do People Drink The Most Beer, Wine And Spirits?" This data was collected by World Health Organisation, Global Information System on Alcohol and Health (GISAH), 2010._

### Type of study 

**What type of study is this (observational/experiment)?**

_This is an observational study._

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

_World Health Organization. (n.d.). Global information system on alcohol and health. World Health Organization. Retrieved October 19, 2021, from https://www.who.int/data/gho/data/themes/global-information-system-on-alcohol-and-health. _


### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

_The dependent variable is alcohol consumption and it it quantitative._

### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**

_The independent variable is country and it is a qualitative._

### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
summary(drinks$country)
```

```{r}
summary(drinks$beer_servings)
```

```{r}
drinks %>% 
ggplot(aes(x = beer_servings)) +
  geom_histogram()
```


```{r}
summary(drinks$spirit_servings)
```

```{r}
drinks %>% 
ggplot(aes(x = spirit_servings)) +
  geom_histogram()
```


```{r}
summary(drinks$wine_servings)
```

```{r}
drinks %>% 
ggplot(aes(x = wine_servings)) +
  geom_histogram()
```


```{r}
summary(drinks$total_litres_of_pure_alcohol)
```

```{r}
drinks %>% 
ggplot(aes(x = total_litres_of_pure_alcohol)) +
  geom_histogram()
```

