---
title: DATA 606 Final Project
author: Leticia Salazar
output:
  html_document:
    theme: flatly
    heightlight: breezedark
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---
$~$

## Introduction

Historically social drinking or moderate alcohol consumption holds an important role in social engagement for many.  According to WHO, World Health Organization, the recorded alcohol consumption per capita (year 15+) in the United States is 47% beer, 18% wine and 35% spirits. 

This raised a question, in general, is there a significant difference in the preferred type of alcohol?

$~$

### Method of data collection
In order to answer my question I collected data from [FiveThirtyEight's](https://fivethirtyeight.com/features/dear-mona-followup-where-do-people-drink-the-most-beer-wine-and-spirits/) article called "Dear Mona Followup: Where Do People Drink The Most Beer, Wine And Spirits?" This data was collected by World Health Organization, Global Information System on Alcohol and Health (GISAH), 2010.

$~$

### Cases 

Each case represents a country around the world along with their beer, spirits and/or wine number of servings, as well as the total liters of pure alcohol. There are 193 total observations in this data set.

$~$

### Type of study 

This is an observational study.

$~$

### Dependent Variable

The dependent variable is alcohol consumption and it it quantitative.

$~$

### Independent Variable

The independent variables are country and types of alcohol and they are qualitative.

$~$

## Data

$~$

```{r}
# load libraries
library(DATA606)
library("tidyverse")
library("tidyr")
library("dplyr")
library("ggplot2")
```

$~$

```{r echo=FALSE}
# load data

drinks <- read.csv("https://raw.githubusercontent.com/letisalba/Data-606/main/Project/drinks.csv", header = TRUE, na.strings = c(""," ", "NA"));
```

$~$

## Exploratory Data Analysis

$~$

```{r}
glimpse(drinks)
```

$~$

```{r}
# Get column names
names(drinks)

# Rename columns
colnames(drinks) <- c("Country", "Beer_Servings", "Spirit_Servings", "Wine_Servings", "Total_Litres_Pure_Alcohol")
head(drinks, n = 5)
```

$~$

```{r}
# commenting out so that it doesn't replicate in my system

#write.csv(drinks, "/Users/letiix3/Desktop/Data-606/Project/csv file/drinks2.csv")
```

$~$

```{r}
summary(drinks)
```

$~$

```{r}

```


## Inference

$~$

**Ho:** countries on the Eastern hemisphere will have less `Total_Litres_Pure_Alcohol` than those in the Western, Northern and Southern hemispheres.

**Ha:** countries on the Eastern hemisphere will not have less `Total_Litres_Pure_Alcohol` than those in the Western, Northern and Southern hemispheres.

$~$

### The Normal Probability Plot

```{r}
# Total_Litres_Pure_Alcohol
qqnorm(drinks$Total_Litres_Pure_Alcohol)
qqline(drinks$Total_Litres_Pure_Alcohol)
```

$~$

```{r}
# Beer_Servings
qqnorm(drinks$Beer_Servings)
qqline(drinks$Beer_Servings)
```

$~$

```{r}
# Spirit_Servings
qqnorm(drinks$Spirit_Servings)
qqline(drinks$Spirit_Servings)
```

$~$

```{r}
# Wine_Servings
qqnorm(drinks$Wine_Servings)
qqline(drinks$Wine_Servings)
```

$~$

### Linear Regression Model
```{r}
# lm for Beer against total litres of pure alcohol
beer_model <- lm(Total_Litres_Pure_Alcohol~ Beer_Servings, data = drinks)
summary(beer_model)
```

$~$

```{r}
anova(beer_model)
```

$~$

```{r}
# lm for Wine against total litres of pure alcohol
wine_model <- lm(Total_Litres_Pure_Alcohol~ Wine_Servings, data = drinks)
summary(wine_model)
```

$~$

```{r}
anova(wine_model)
```

$~$

```{r}
# lm for Spirit against total litres of pure alcohol
spirit_model <- lm(Total_Litres_Pure_Alcohol~ Spirit_Servings, data = drinks)
summary(spirit_model)
```

$~$

```{r}
anova(spirit_model)
```

$~$

## Plots

$~$

```{r}
# Beer servings by Country; top 40
drinks %>% 
  arrange(desc(Beer_Servings)) %>% 
  head(40) %>% 
  ggplot(aes(y = reorder(Country, Beer_Servings),
             x = Beer_Servings,
                  fill= Country))+
  geom_col()+
  xlab("Beer Servings") +
  ylab("Country") +
  theme(legend.position = "none")
```

$~$

```{r}
# Spirits serving by Country; top 40
drinks %>% 
  arrange(desc(Spirit_Servings)) %>% 
  head(40) %>% 
  ggplot(aes(y = reorder(Country, Spirit_Servings),
             x = Spirit_Servings,
                  fill= Country))+
  geom_col()+
  xlab("Spirit Servings") +
  ylab("Country") +
  theme(legend.position = "none")
```

$~$

```{r}
# Wine servings by Country; top 40
drinks %>% 
  arrange(desc(Wine_Servings)) %>% 
  head(40) %>% 
  ggplot(aes(y = reorder(Country, Wine_Servings),
             x = Wine_Servings,
                  fill= Country))+
  geom_col()+
  xlab("Wine Servings") +
  ylab("Country") +
  theme(legend.position = "none")
```

$~$

```{r}
# Total alcohol in litres by Country; top 40
drinks %>% 
  arrange(desc(Total_Litres_Pure_Alcohol)) %>% 
  head(40) %>% 
  ggplot(aes(y = reorder(Country, Total_Litres_Pure_Alcohol),
             x = Total_Litres_Pure_Alcohol,
                  fill= Country))+
  geom_col() +
  xlab("Total Litres of Pure Alcohol") +
  ylab("Country") +
  theme(legend.position = "none")
```

$~$

```{r}
drinks %>% 
ggplot(aes(x = Beer_Servings)) +
  geom_histogram()
```

$~$

```{r}
summary(drinks$Spirit_Servings)
```

$~$

```{r}
drinks %>% 
ggplot(aes(x = Spirit_Servings)) +
  geom_histogram()
```

$~$

```{r}
summary(drinks$Wine_Servings)
```

$~$

```{r}
drinks %>% 
ggplot(aes(x = Wine_Servings)) +
  geom_histogram()
```

$~$

```{r}
summary(drinks$Total_Litres_Pure_Alcohol)
```

$~$

```{r}
drinks %>% 
ggplot(aes(x = Total_Litres_Pure_Alcohol)) +
  geom_histogram()
```

$~$

```{r}

```

$~$

## Conclusion

$~$

## References

$~$

World Health Organization. (n.d.). Global information system on alcohol and health. World Health Organization. Retrieved October 19, 2021, from https://www.who.int/data/gho/data/themes/global-information-system-on-alcohol-and-health. 

$~$

Ritchie, H., &amp; Roser, M. (2018, April 16). Alcohol consumption. Our World in Data. Retrieved November 15, 2021, from https://ourworldindata.org/alcohol-consumption. 

$~$

WHO. (2016). 19089 United States of america - world health organization. Retrieved December 5, 2021, from https://www.who.int/substance_abuse/publications/global_alcohol_report/profiles/usa.pdf.