---
title: "DATA 606 Data Project Proposal"
author: "Leticia Salazar"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: yeti
    highlight: tango
    toc: yes
    toc_float: yes
---
$~$

### Data Preparation
```{r}
# load packages / libraries
library("tidyverse")
library("tidyr")
library("dplyr")
library("ggplot2")
```
$~$

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data
drinks <- read.csv("https://raw.githubusercontent.com/letisalba/Data-606/main/Project/drinks.csv", header = TRUE, na.strings = c(""," ", "NA"));
```
$~$
```{r}
glimpse(drinks)
```
$~$
```{r}
# Get column names
names(drinks)
```
$~$
```{r}
# Rename columns
colnames(drinks) <- c("Country", "Beer_Servings", "Spirit_Servings", "Wine_Servings", "Total_Litres_Pure_Alcohol")
```
$~$

### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your data set allows for.**

_What country has the highest alcohol consumption?_

$~$

### Cases 

**What are the cases, and how many are there?**

_Each case represents a country around the world along with their beer, spirits and/or wine number of servings, as well as the total liters of pure alcohol. There are 193 total observations in this data set._

$~$

### Data collection 

**Describe the method of data collection.**

_The data was collected from [FiveThirtyEight's](https://fivethirtyeight.com/features/dear-mona-followup-where-do-people-drink-the-most-beer-wine-and-spirits/) article called "Dear Mona Followup: Where Do People Drink The Most Beer, Wine And Spirits?" This data was collected by World Health Organisation, Global Information System on Alcohol and Health (GISAH), 2010._

$~$

### Type of study 

**What type of study is this (observational/experiment)?**

_This is an observational study._

$~$

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

_World Health Organization. (n.d.). Global information system on alcohol and health. World Health Organization. Retrieved October 19, 2021, from https://www.who.int/data/gho/data/themes/global-information-system-on-alcohol-and-health. _

$~$

### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

_The dependent variable is alcohol consumption and it it quantitative._

$~$

### Independent Variable

**What is the independent variable? Is it quantitative or qualitative?**

_The independent variable is country and it is a qualitative._

$~$

### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
summary(drinks$Country)
```



$~$
```{r}
summary(drinks$Beer_Servings)
```
$~$


$~$

```{r}
drinks %>% 
ggplot(aes(x = Beer_Servings)) +
  geom_histogram()
```
$~$

```{r}
summary(drinks$Spirit_Servings)
```
$~$
```{r}
drinks %>% 
ggplot(aes(x = Spirit_Servings)) +
  geom_histogram()
```
$~$

```{r}
summary(drinks$Wine_Servings)
```
$~$
```{r}
drinks %>% 
ggplot(aes(x = Wine_Servings)) +
  geom_histogram()
```
$~$

```{r}
summary(drinks$Total_Litres_Pure_Alcohol)
```
$~$
```{r}
drinks %>% 
ggplot(aes(x = Total_Litres_Pure_Alcohol)) +
  geom_histogram()
```

